```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

### Part 1 : Data Preparation

In this first part, we will first download the data into R. Then, we will apply the necessary transformations on these data so that they can be correctly exploited for the following analysis. 

## Import dataset 
```{r}
Mountain_data <- read_excel("../data/1-s2.0-S2352340919311710-mmc1.xlsx", sheet = "FieldDATA")
Mountain_data$Date <- as.Date(Mountain_data$Date, "%d/%m/%Y")

head(Mountain_data)
```

We are deleting the variables that are not useful for use. The variables such as the **UTM grid zone**, **UTM_X** and **UTM_Y** are related to the samples and we are not wanting to make predictions based on those characteristics but rather on the actual soil composition. 
The variable **Elevation**, **slope**, **plot_ID**, **Orientation** will not help us neither because it is related to the place where the sample has been taken. It is a certain amount above the some tree limits.
We rename the variable **Muntain Range** as **Mountain_range** to facilitate future use. 

After these changes, he database *Mountain_data* is composed of 430 observation and 34 variables.

- **Country** : *chr* Spain, Chile
- **Mountain_range** : *chr* Sierra de Guadarrama, Central Pyrenees
- **Locality** : *chr* Locality of the sample (there is different location on the same mountain)
- **Plot** : *num* ?
- **Subplot** : *num* ?
- **Date** : *Date* Date of the sample
- **Day** : *num* Day of the sample
- **Month** : *num* Month of the sample
- **Year** : *num* Year of the sample
- **Radiation** : *num* Potential solar radiation (0-1)
- **Phos_P** : *num* Phosphatase enzyme.  Measured on dry soil sampled under plant (µmol/gr dry soil/h)
- **Glu_P** : *num* β-glucosidase enzyme.  Measured on dry soil sampled under plant canopy (µmol/gr dry soil/h)
- **SOC_P** : *num* soil organic carbon. Measured on dry soil sampled under plant canopy (%)
- **NT_P** : *num* soil total nitrogen. Measured on dry soil sampled under plant canopy (mg/g soil)
- **PT_P** : *num* available phosphorus. Measured on dry soil sampled under plant canopy (mg/g soil)
- **K_P** : *num* potassium content. Measured on dry soil sampled under plant canopy (mg/g soil)
- **pH_P** : *num* pH. Measured on dry soil sampled under plant canopy 
- **Cond_P** : *num* electrical conductivity. Measured on dry soil sampled under plant canopy (µS/cm)
- **Phos_B** : *num* Phosphatase enzyme.  Measured on dry soil sampled from bare soil areas (µmol/gr dry soil/h)
- **Glu_B** : *num* β-glucosidase enzyme.  Measured on dry soil sampled from bare soil areas (µmol/gr dry soil/h)
- **SOC_B** : *num* soil organic carbon. Measured on dry soil sampled from bare soil areas (%)
- **NT_B** : *num* soil total nitrogen. Measured on dry soil sampled from bare soil areas (mg/g soil)
- **PT_B** : *num* available phosphorus. Measured on dry soil sampled from bare soil areas (mg/g soil)
- **K_B** : *num* potassium content. Measured on dry soil sampled from bare soil areas (mg/g soil)
- **pH_B** : *num* pH. Measured on dry soil sampled from bare soil areas
- **Cond_B** : *num* electrical conductivity. Measured on dry soil sampled from bare soil areas (µS/cm)
- **Phos_T** : *num* Phosphatase enzyme. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (µmol/gr dry soil/h)
- **Glu_T** : *num* β-glucosidase enzyme.  Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (µmol/gr dry soil/h)
- **SOC_T** : *num* soil organic carbon. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (%)
- **NT_T** : *num* soil total nitrogen. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (mg/g soil)
- **PT_T** : *num* available phosphorus. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (mg/g soil)
- **K_T** : *num* potassium content. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (mg/g soil)
- **pH_T** : *num* pH. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site
- **Cond_T** : *num* electrical conductivity. Values averaged and weighted by the respective cover of bare ground and vegetated area in each site (µS/cm)


```{r}

Mountain_data <- Mountain_data[, c(2:5, 7:11, 18:42)]
colnames(Mountain_data)[2] <- "Mountain_range"
```

Now that we have the variables we want to work with, we see that they are all considered as numeric. However, we know that some of them should be categorical as they have different levels. We are going to transform **Country**, **Mountain_Range**, **Locality**, **Plot** and **Subplot** as factors. 

```{r}
Mountain_data$Country <- as.factor(Mountain_data$Country)
Mountain_data$Mountain_range <- as.factor(Mountain_data$Mountain_range)
Mountain_data$Locality <- as.factor(Mountain_data$Locality)
Mountain_data$Plot <- as.factor(Mountain_data$Plot)
Mountain_data$Subplot <- as.factor(Mountain_data$Subplot)
```

We can export this cleaned dataset on the _data_ folder of our project. 

```{r}
write.csv(Mountain_data,"../data/Mountain_data_cleaned.csv", row.names = FALSE)
```





