
```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

### part 2 : Exploratory Data Analysis 

This section is dedicated to understanding the data. We will provide an analysis of the data set using a visual approach in order to summarize their main characteristics. 

Let's import the cleaned dataset that we created. 

```{r}
Mountain_data_cleaned <- read.csv("../data/Mountain_data_cleaned.csv")
str(Mountain_data_cleaned)
```

We have to transform the class of the variables **Country**, **Mountain_Range**, **Locality**, **Plot** and **Subplot** because they are categorical variables. So we will transform them as factors. 

```{r}
Mountain_data_cleaned$Country <- as.factor(Mountain_data_cleaned$Country)
Mountain_data_cleaned$Mountain_range <- as.factor(Mountain_data_cleaned$Mountain_range)
Mountain_data_cleaned$Locality <- as.factor(Mountain_data_cleaned$Locality)
Mountain_data_cleaned$Plot <- as.factor(Mountain_data_cleaned$Plot)
Mountain_data_cleaned$Subplot <- as.factor(Mountain_data_cleaned$Subplot)

Mountain_data_cleaned$Date <- as.Date(Mountain_data_cleaned$Date)

summary(Mountain_data_cleaned)
```

There are 2 missing values in the feature **Glu_P**. One is at the instance number 377 and the other one is at the instance 378. 

We have more observations about the mountain "Sierra de Guadarrama" (195) compared to "Central Andes" (100) and "Central Pyrenees" (135). As the differences between the number of observation is big enough for us to be careful on the results and consider to balance the data if it is needed. 

```{r}
which(is.na(Mountain_data_cleaned$Glu_P))
```

```{r}
dfSummary(Mountain_data_cleaned, style="grid")
```


```{r}
describe(Mountain_data_cleaned)
introduce(Mountain_data_cleaned)
plot_intro(Mountain_data_cleaned)
```

## Data visualization : Plotting the data

We plot the numerical variables. 

```{r}
plot_histogram(Mountain_data_cleaned)
plot_density(Mountain_data_cleaned)
```
Many variables appear with a distribution that looks like the log-normal distribution.

+ The variables **pH_B** and **pH_T** are not normally distributed. 
+ Most of the variables are right-tailed. 


```{r}
plot_boxplot(Mountain_data_cleaned, by = 'Mountain_range',  ncol = 2,
             title = "Side-by-side boxplots")
```

We can see the real differences between the mountains. 

+ Generally speaking, it seems that the mountain "Sierra de Guadarrama" has a higher value of **Glu_P**, **Phos_P**, **SOC_P**, **Glu_B**, **Phos_B**, **SOC_B**, **Glu_T**, **Phos_T** and **SOC_T**.
+ There are a lot of outliers in almost all the features. 

We plot the categorical variables. 

```{r}
Mountain_data_cleaned %>%
  keep(is.factor) %>%
  gather() %>%
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_bar(stat="count") +
  coord_flip()
```


```{r}
ggplot(Mountain_data_cleaned) +
  geom_bar(aes(x = Mountain_range))
```
Can comment on the number of different observations and the effect on accuracy. **Can focus more on sensibility and sensitivy** if needed because indeed there is twice more informations on Sierra de Guadarrama.

As described above with the summary output of the data, we see that we have more information on the mountain "Sierra de Guadarrama". There is twice more information compared to the mountain "Central Andes". Our final result might be affected on a bad way because the model will tend to produce a good accuracy (so having a tendency to predict "Sierra de Guadarrama" more often) but it will not be good enough to predict a new instance. 

We will have to see if we will need to balance our data to get a better model. 


Correlation plot between continuous variables : 
```{r}
plot_correlation(Mountain_data_cleaned, type= 'c', cor_args = list( 'use' = 'complete.obs'))
```

From the correlation plot it seems that some pattern can be observed. The variables concerning the **Phosphatase enzyme** seems to be positively correlated with the variable about **Soil organic carbon**. 

```{r}
Mountain_data_cleaned %>%
  select(Phos_P, Phos_B, Phos_T, SOC_P, SOC_B, SOC_T) %>%
  ggpairs()
```


Make a regression Mountain_range ~ ... to see the relevant variables 
+
Make a PCA


## Principal Component Analysis (PCA) exploration

- Select only the numerical values:

```{r}
Mountain_data.num <- select_if(Mountain_data_cleaned, is.numeric)
to.delet <- c('Day', 'Month', 'Year')
Mountain_data.num <- select(Mountain_data.num, -to.delet)

```


```{r}
Mountain_data.pca <- prcomp(na.omit(Mountain_data.num), center = TRUE)
summary(Mountain_data.pca)
```


Here, as the command _prcomp_ do not allow NAs in the data. We use the command _na.omit_ on our reduced data containing the numerical values to omit all NAs cases from the data frame.


```{r}
resultats_pca <- PCA(Mountain_data.num, scale.unit = TRUE, ncp = 5, graph = FALSE)
resultats_pca
```

```{r}
eig.val <- get_eigenvalue(resultats_pca)
eig.val
```

```{r}
fviz_eig(resultats_pca, addlabels = TRUE)
```

```{r}
fviz_pca_var(resultats_pca)
```

```{r}
var <- get_pca_var(resultats_pca)
var
```

```{r}
corrplot(var$cos2, is.corr = FALSE)
```

```{r}
fviz_pca_ind(resultats_pca)
```


```{r}
fviz_pca_biplot(resultats_pca, repel = TRUE,
                col.var = "blue", # Couleur des variables
                col.ind = "red"  # Couleur des individues
                )
```


